An analysis of PCA image reconstruction for classification

### 2010-04-13 paper/paper.tex
Seção de experimentos finalizada.

However, this does not mean that all the four components have the same relevance to classification. In the proposed classifier, each component contributes equally to the total error, but our experiments show that this relation is not true. Some components have more importance than others and therefore have to be adjusted accordingly. It is also worth noting that the classifier can be made more flexible if we introduce a threshold parameter and compare the classification score (i.e. the total error) agains it, instead of keep it always fixed to zero. In this way, the user can adjust the classifier specificity, depending on his particular application.

@@

Again, we plot the reconstruction error distributions, but now segregating error from grayscale PCs and from edges PCs. These plots are shown in Figure XX, that also displays a plot of the combined errors distribution (i.e. grayscale and edge errors summed). At this time, we get a different scenario. All the three plots exhibit a clear boundary that separate samples from different classes. This indicated that there is a lot of redundancy between errors from grayscale PCs and from edge PCs and one of them can be safely removed from the classifier with a minor impact on classification accuracy. Our results indicate that the errors from grayscale PCs generally yields a slightly worse result, as shown in the Table XX. Therefore, it is possible to compose a classifier only with the edge errors and speed up the classification significantly (since two reconstructions will not have to be computed anymore) losing less than $1\%$ of the classification accuracy. Actually, the classifier based only on the reconstruction errors from the edge PCs sometimes performed even better in our experiments (when $k >= 200$), as also shown in Table XX.

@@

% From the description above, we notice that each reconstruction contributes equally to the total error. Therefore, it is assumed that each one has the same relevance to the classification, what, as our experiments show, it is not the case. We found that some reconstructions contribute more to the classification and it is possible to improve the classification accuracy by adjusting their weights when computing the total error.
% but, as our experiments show, this is not the case. It seems 
% To have a better understanding on how this classifier works, we

>>>>

However, this does not mean that all the four components have the same relevance to classification. In the proposed classifier, each component contributes equally to the total error, but our experiments show that this relation is not true. Some components have more importance than others and therefore have to be adjusted accordingly. So, we propose the \emph{weighted total reconstruction error}, given by

\begin{equation}
  d_w = w_{gn} d_{gn} + w_{en} d_{en} - w_{gp} d_{gp} - w_{ep} d_{ep}
  \label{weighted_total_error}
\end{equation}

\noindent where $w_{gn}$, $w_{en}$, $w_{gp}$, $w_{ep}$ are weights that adjust the relevance of each error to the classification (more relevant errors must have larger weights). They can be set manually or be determined by some optimization algorithm in relation to the training set. We use a genetic algorithm in our experiments.

It is also worth noting that the classifier can be made more flexible if we introduce a threshold parameter and compare the classification score (i.e. the total error) agains it, instead of keep it always fixed to zero. In this way, the user can adjust the classifier specificity, depending on his particular application.

@@

Again, we plot the reconstruction error distributions, but now segregating error from grayscale PCs and from edges PCs. These plots are shown in Figure XX, that also displays a plot of the combined errors distribution (i.e. grayscale and edge errors summed). At this time, we get a different scenario. All the three plots exhibit a clear boundary that separate samples from different classes. This indicate that there is a lot of redundancy between errors from grayscale PCs and from edge PCs and one of them can be safely removed from the classifier with just a minor impact on classification accuracy. Our results indicate that the errors from grayscale PCs generally yields a slightly worse result, as shown in the Table XX. Therefore, it is possible to compose a classifier only with the edge errors and speed up the classification significantly (since two reconstructions will not have to be computed anymore). In general, such classifier drops the classification accuracy in less than $1\%$ when compared with a classifier that uses the total error. Actually, the classifier based only on the edge errors sometimes performs even better in our experiments (when $k >= 200$), as also shown in Table XX.

@@

% What reference use to GA?
These findings also point out that, among all the four reconstruction errors, some are more relevant to the classification than others. However, on the total reconstruction error (equation \ref{total_error}) each error contributes equally. So, we use the weighted total reconstruction error (equation \ref{weighted_total_error}) to improve the classification accuracy. The problem with using weights is how to choose them. We can try some values and make adjusts manually, test after test. But we can also employ an optimization algorithm that looks for the values in an automated way. In this work, we use a genetic algorithm \ref{ga}. It initially chooses a set of random values and uses operations as crossover, mutation and selection to improve them during many iterations. These operations are inspired by evolutionary biology; hence the name genetic algorithms. The values are adjusted in relation to a fitness function. In our case, this functions aims to correctly classify the largest number of training samples using a given set of weights.

As this algorithm performs a local optimization and therefore is not optimum, we have run it five times for each value of $k$. We then have chosen the set of weights that gets the best classification rates among these five; they are listed on Table XX. Finally, we classify the test samples considering the weighted total reconstruction error. The produced ROC curves always have a larger AUC than the corresponding versions without weights for all values of $k$ we tested. This is shown in Figure XX. The difference is more pronounced for large values of $k$ and becomes small when it diminishes. Still, the AUC of the first classifier is always larger, what indicates that the weighted error is a better classification criteria than the non-weighted one. The chart of Figure XX also shows that the AUC of the weighted classifier is more stable when $k$ varies.

% The better value of k depends on the particular application

% k pgp pep pgn pen Total AUC Weighted AUC
% 400 0.4456  0.5472  0.2785  0.9340  0.9744  0.9981
% 300 1.6523  0.7827  0.9183  2.2065  0.9826  0.9970
% 200 0.3101  0.6551  0.1338  0.9962  0.9877  0.9981
% 100 0.8472  1.3322  0.7671  1.5308  0.9948  0.9990
% 50  0.2236  0.8294  0.1905  0.9043  0.9972  0.9991
% 25  0.9172  1.1696  0.7768  1.4662  0.9984  0.9989

% TODO Figure with k vs AUC for non-weighted and weighted
